---
title: "Importing Data"
author: "Patrick Mathias"
output:
  pdf_document: default
  html_document: default
---

## Your Turn 1

The first chunk in the document is typically where you load all packages you will use throughout the document.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(janitor)
library(knitr) # making nice tables
```

## Where am I?

Use `getwd()` to show your working directory.

```{r wd}
getwd()
```

## Load a file

```{r load_file}
catalog <- read_csv("data/test_catalog.csv")
```

What does RStudio tell you after you import the data?

## Your Turn 2

1. Locate and load the "orders_data_set.csv" file from the data folder into an object (data frame) called "orders". 
2. Within your Environment tab (top right), click the name of the object and review the raw data.
3. Summarize the data using the `summary()` function.

```{r}
orders <- 
summary( )
```

## Parsing data on import

The readr and readxl packages generally are effective in identifying the data type of a column when it is being imported. They will automatically parse the data with the correct type most of the time. For example, if a column has only whole numbers, the data frame that you import the data into will have a integer data type for that column. The same is true for dates and times, but there are a variety of formats that the import functions don't recognize unless you provide a date and/or time format.

There are additional arguments to readr functions that can explicitly specify the data type for columns.

```{r}
orders <- read_csv("data/orders_data_set.csv", 
                   col_types = cols(
                     order_time = col_datetime("%m/%d/%Y %H:%M")
                     )
                   )
summary(orders)
```

## Your Turn 3

Import the result_time and review_time columns with the datetime data type and confirm the data types are correct.

```{r}
orders <- 
summary(orders)
```


## Tabulating

`tabyl` function can create quick cross tabluations to drill down into categorical fields.

```{r}
tabyl(orders, department)
```


```{r}
tabyl(orders, department, order_status_c_descr)
```

## Making your tables pretty

```{r}
kable(tabyl(orders, department, pref_list_type))
```


