---
title: "Transforming Data"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Your turn 1

```{r setup}
library(tidyverse)  # Provides functions used throughout this session
library(readxl) # Provides function for reading in excel workbooks
orders <- read_excel("/cloud/project/data/orders_data_set.xlsx")
```

## Orders

```{r}
orders
```

```{r}
summary(orders)
```


## Select

```{r}
select(orders, description, department)
select(orders, c(1,4))
```

## Filter

```{r}
filter(orders, patient_id == 508061)
```

```{r}
filter(orders, proc_code == "BMP")
```

## Your turn 2

Use the 2 code blocks below, and `filter()`, to show how you would filter the `orders` dataset.

Modify the code in this block to find:

a. Every order_id that is greater than 100000

```{r, eval = FALSE}
orders_2a <- filter(orders, ________)
orders_2a
```

After modifying the code above, copy and paste it into the next two code blocks, and alter it to show:

b. All of the orders where lab_status_c_descr is equal to "Final result" (assign this to orders_3b)

```{r}

```

## Arrange

```{r}
arrange(orders, patient_id)
```

## Steps and the pipe

```{r}
VITD <- orders %>%
		 filter(description == "1,25 DIHYDROXY VITAMIN D") %>%
		 select(department, ordering_route, pref_list_type) %>%
		 arrange(department)
VITD
```

## Your turn 3

Use `%>%` to write a sequence of functions that: 

1. Filters to orders coming from the `department` "BEHAVIORAL HEALTH CLINIC" (case-sensitive)
2. Selects the `description`, `ordering_route`, and `pref_list_type` columns  
3. Arrange the dataset by the `description` and `ordering_route` column.

Assign the output to a new variable (orders_6).

CHALLENGE: Use your mouse to select the name of the new data frame from the list of data sets in the upper-right pane of Rstudio. Do you notice any interesting test ordering patterns?

```{r, eval = FALSE}
orders_3 <- orders %>% 
  filter() %>% 
  select() %>% 
  arrange()
```

## mutate

```{r}
orders %>%
	mutate(coded_order_id = order_id/2)
```

## Your turn 4

The weekdays() function will return the weekday for any date.

1. Use the weekdays() function to make a new column which contains the day of the week that each order was placed.
2. Then select this new column and the order_time column.
3. Assign the result to a new variable orders_7

```{r, eval = FALSE}
orders_4 <- orders %>%
  
  
```

## Replace a column with `mutate()`

```{r}
orders %>%
	mutate(order_id = as.character(order_id))
```

## Conditionally replace values in a column with `mutate()`

```{r}
orders %>%
	mutate(proc_code = ifelse(proc_code %in% c("CBC","CBD"), 
	                          "CBC", 
	                          proc_code)) 
```

## Using `case_when()` with `mutate()`

In some cases we want to rename categories of data of apply multiple logic conditions to categorize an observation. `case_when()` can be used along with `mutate()` to provide a list of conditions.

```{r}
orders <- orders %>%
  mutate(pref_list_cat = case_when(
    pref_list_type == "Clinic Preference List" ~ "clinic",
    pref_list_type == "Provider Preference List" ~ "provider",
    pref_list_type == "None" ~ NA_character_
  ))
```

# Take aways

* Extract variables with `select()`  
* Extract cases with `filter()`  
* Arrange cases, with `arrange()`  

* Make new variables, with `mutate()`  

* Connect operations with `%>%`  

