---
title: "Summarizing Data"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Setup

```{r setup}
library(tidyverse) 
library(readxl) 

orders <- read_csv("../data/orders_data_set.csv", 
                   col_types = cols(
                     order_time = col_datetime("%m/%d/%Y %H:%M"),
                     result_time = col_datetime("%m/%d/%Y %H:%M")  # Specifying equivalent date format to variables used below
                     )
                   )
```

## Summarize

```{r}
orders %>%
  select(order_id, patient_id) %>%
  head(4) %>%
	summarize(order_count = n(),
	          pt_count    = n_distinct(patient_id)) 
```

## Your Turn 1

Add onto the code in the above chunk to calculate the items for a and b below.

*Hints:* 

- Start by adding a new code block under each item (copy/paste above or CTRL-ALT-I).
- Then copy code from the previous code block and edit to determine the mean number of orders per patient. 
- Next, add code to calculate the mean count of orders per department. 
- Assign the results to new objects, orders_1a and orders_1b.

a) Mean count of orders per patient
 
```{r}
# solution
orders_1a <- orders %>%
	summarize(order_count = n(),
	          pt_count = n_distinct(patient_id),
	          pt_order_count_mean = n() / n_distinct(patient_id))
orders_1a
```

b) Mean count of orders per department

```{r}
# solution
orders_1b <- orders %>%
	summarize(order_count = n(),
	          dept_count = n_distinct(department),
	          dept_order_count_mean = order_count / dept_count)
```


## Your Turn 2

Use summarize() to calculate:

a) The date of the first (or minimum) order

```{r, eval=FALSE}
orders %>% 
	summarize(______________)
```

```{r}
#solution
orders_2a <- orders %>% 
	summarize(order_time_min = min(order_time))
orders_2a
```

b) The median time difference between order_time and result_time

*Hint: Refer to function help for missing data (NA) handling*

```{r, eval=FALSE}
orders %>%
  mutate(result_interval = ______________) %>%
  summarize(result_interval_med = __________________________)
```


```{r}
#solution
orders_2b <- orders %>% 
	mutate(result_interval = result_time - order_time) %>%
	summarize(result_interval_med = median(result_interval, na.rm = T))
orders_2b
```

## group_by

```{r}
orders %>% 
  group_by(patient_id, department) %>%
  head(4)
```

## group_by %>% summarize

```{r}
orders %>%
	group_by(patient_id) %>%
	summarize(order_count = n()) %>%
	summarize(pt_order_count_mean = mean(order_count))
```

## Your Turn 4

a) Calculate the median number of orders per patient

```{r}
# solution
orders_4a <- orders %>%
	group_by(patient_id) %>%
	summarize(order_count = n()) %>%
	summarize(pt_order_count_median = median(order_count))
orders_4a
```

b) Calculate the maximum number of TSH orders (`proc_code` == "TSH") per patient

```{r, eval=FALSE}
orders_4b <- orders %>%  
	______(_______ == ______) %>%	
	group_by(patient_id) %>%
	summarize(order_count = n()) %>%
	summarize(pt_order_count_max = _____________)
orders_4b
```

```{r}
#solution
orders_4b <- orders %>%  
	filter(proc_code == "TSH") %>%	
	group_by(patient_id) %>%
	summarize(order_count = n()) %>%
	summarize(pt_order_count_max = max(order_count))
orders_4b
```

c) (*Bonus*) The 5th and 95th percentile of the number of orders per patient

```{r}
#solution
orders_4c <- orders %>%
	group_by(patient_id) %>%
	summarize(order_count = n()) %>%
	summarize(pt_order_count_q05 = quantile(order_count, probs=0.05),  
	          pt_order_count_q95 = quantile(order_count, probs=0.95))
orders_4c
```

## Your Turn 5

Calculate the mean order count per patient for each department

*Hint: summarize() rolls up a single grouping variable at a time*

```{r, eval=FALSE}
orders_5 <- orders %>%
  ________(______________) %>%
  summarize(_______________) %>%
  summarize(_______________________)
```


```{r}
#solution
orders_5 <- orders %>%	    
	group_by(department, patient_id) %>%	    
	summarize(order_count = n()) %>%    # summarize over department-patient_id groups
	summarize(pt_order_count_mean = mean(order_count)) # summarize over department groups
orders_5
```
