---
title: "Statistics"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Setup

```{r setup}
library(tidyverse) 
library(readxl) 

orders <- read_csv("../data/orders_data_set.csv", 
                   col_types = cols(
                     order_time = col_datetime("%m/%d/%Y %H:%M"),
                     result_time = col_datetime("%m/%d/%Y %H:%M")  # Specifying equivalent date format to variables used below
                     )
                   )
```

## Compare two sets of count data

- Is the distribution of the count of orders per patient different in the Internal Medicine and Family Medicine clinics?

# Describe central tendency of order counts for each department

```{r}
orders %>%	    
	group_by(department, patient_id) %>%	    
	summarize(order_count = n()) %>%
	summarize(order_count_median = median(order_count), 
		   order_count_mean = mean(order_count)) %>%
	filter(department %in% c("INTERNAL MEDICINE CLINIC", 
					"FAMILY MEDICINE CLINIC"))
```


## Prepare dataset with order counts per patient for each department

```{r}
orders_per_pt_dept <- orders %>%	    
	group_by(department, patient_id) %>%	    
	summarize(order_count = n()) %>%  # summarize over department-patient_id groups
  ungroup() %>%
	filter(department %in% c("INTERNAL MEDICINE CLINIC", 
					"FAMILY MEDICINE CLINIC"))

orders_per_pt_dept
```

## Visualize the distribution of order counts

```{r}
orders_per_pt_dept %>%	        
        ggplot() +
	      geom_histogram(aes(x = order_count), binwidth = 1) +
    facet_wrap(~department)
```



## Compare the distributions of order counts

```{r}
wilcox.test(order_count ~ department, 
            data        = orders_per_pt_dept, 
            alternative = "two.sided", 
            paired      = FALSE, 
            conf.int    = TRUE)
```

## What Else?

## Can we predict the number of orders for each department in September?

## Prepare data for regression

```{r}
library(lubridate)
orders_per_dept_month <- orders %>%
	mutate(order_month = month(order_time, label=TRUE, abbr=TRUE)) %>%	 # Assign month
	group_by(department, order_month) %>%	    
	summarize(order_count = n()) %>%
	ungroup() %>%
  spread(key = order_month, value = order_count) # Spread Months from values of a variable to separate variables
```

## Your Turn 1

a) Predict October order counts for each department using September order counts
```{r, eval=FALSE}
model <- lm(____ ~ ______, 
                data = ____________)
summary(model)
coef(model)
```

```{r}
#solution
model <- lm(Oct ~ Sep, 
            data = orders_per_dept_month)
summary(model)
coef(model)

ggplot(data = orders_per_dept_month) +
  geom_point(aes(x=Sep, y=Oct)) +
  geom_smooth(method="lm", aes(x=Sep, y=Oct))
```
b) Use deming (orthogonal) regression to compare monthly order counts

```{r, eval=FALSE}
_________
model_deming <- _________
printSummary(model_deming)
model_deming@glob.coef
```

```{r}
#solution
# install.package("mcr")
library(mcr)
model_deming <- mcreg(orders_per_dept_month$Sep, orders_per_dept_month$Oct, 
               method.reg = "Deming") 
printSummary(model_deming)
getCoefficients(model_deming)

ggplot(data = orders_per_dept_month) +
  geom_point(aes(x=Sep, y=Oct)) +
  geom_smooth(method="lm", aes(x=Sep, y=Oct), se=FALSE) +
  geom_abline(intercept = getCoefficients(model_deming)[1],
              slope = getCoefficients(model_deming)[2],
              color="red")
```

c) Try to improve upon the linear regression model by adding more predictors to the right-hand side of the regression formula

```{r, eval=FALSE}
model2 <- lm(____ ~ ______ + _______, 
   data = ____________)
summary(model2)
```

```{r}
#solution
model2 <- lm(Oct ~ Sep + Aug, 
   data = orders_per_dept_month)
summary(model2)

model3 <- lm(Oct ~ Sep + Aug + Nov, 
   data = orders_per_dept_month)
summary(model3)

anova(model, model2, model3) # Compare models
```